---
title: "lab03"
author: "Kaila An"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1. Read in the data
First download and then read in with data.table:fread()

```{r}
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
met <- data.table::fread("met_all.gz")
```


## Step 2. Check the dimensions, headers, footers. How many columns, rows are there? 

```{r}
dim(met)
head(met)
tail(met)
```

## Step 3. Take a look at the variables.

```{r}
str(met)
```

## Step 4. Take a closer look at the key variables.

```{r}
table(met$day)
table(met$year)
table(met$hour)
```


```{r}
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
```


Replace elevations with 9999 as NA.


```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```


The weather station with the highest elevation is at `r max(met$elev, na.rm=TRUE)` meters.

```{r}
met <- met[temp>-15][order(temp)]
```

## Step 5. Check the data against an external data source.

## Step 6. Calculate summary statistics.

```{r}
met[elev==max(elev,na.rm=TRUE), summary(wind.sp)]
met[elev==max(elev,na.rm=TRUE), summary(temp)]
```


```{r}
met[elev==max(elev,na.rm=TRUE), .(
  temp_wind     = cor(temp,wind.sp,  use = "complete"),
  temp_hour     = cor(temp,hour,     use = "complete"),
  wind.sp_day   = cor(wind.sp,day,   use = "complete"),
  wind.sp_hour  = cor(wind.sp,hour,  use = "complete")
)]

```

# Step 7. Exploratory Graphs

Look at key variables

```{r}
hist(met$elev)
```

```{r}
hist(met$temp)
```

```{r}
hist(met$wind.sp)
```

```{r}
if (!require(leaflet)) {
  install.packages("leaflet")
  library(leaflet)
}

elev <- met[elev==max(elev)]
```

```{r}
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

The time series of temperature and wind speed at this location.

```{r}
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
```

```{r}
elev <- elev[order(date)]
head(elev)
```

```{r}
plot(elev$date, elev$temp, type='l')
```

The temperature reached a minimum of 0.5 and a maximum of 15.  Significantly lower temperatures were experienced between Aug 07 and Aug 12 and the higher temperatures were felt between Aug 18 to Aug 22.

```{r}
plot(elev$date, elev$wind.sp, type='l')
```

Wind speeds were lower between Aug 02 and Aug 07, as well as Aug 20 to Aug 22.  The higher wind speeds were felt between Aug 14 and Aug 17 and Aug 23 and Aug 26.













