---
title: "lab04"
author: "Kaila An"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-libraries}
library(lubridate)
library(tidyverse)
library(data.table)
```



## Step 1. Read in the data
First download and then read in with data.table:fread()

```{r read-data, cache=TRUE}
if (!file.exists("../lab04/met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("../lab04/met_all.gz")
```


## Step 2. Prep Data

Remove temperatures less than -17C and change elev 9999 to missing value code.
```{r remove-low-temp}
met <- met[ temp > -17][elev == 9999.0 , elev := NA]
```

#check no 9999s in other important variables

Generate a date variable using the functions as.Date() (hint: You will need the following to create a date paste(year, month, day, sep = "-")).
```{r create-ymd, cache=TRUE}
met <- met[ , ymd := as.Date(paste(year, month, day, sep = "-"))] 
```


Using the data.table::week function, keep the observations of the first week of the month.

```{r}
met[, table(week(ymd))]
met <- met[ week(ymd) == 31 ]
```


Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.

```{r}
met[, .(
  temp          = max(temp,na.rm=T),
  rh            = max(rh,na.rm=T),
  wind.sp       = max(wind.sp,na.rm=T),
  vis.dist      = max(vis.dist,na.rm=T),
  dew.point     = max(dew.point,na.rm=T),
  lat           = max(lat,na.rm=T),
  lon           = max(lon,na.rm=T),
  elev          = max(elev,na.rm=T)
)]
```
No more 9999s in our dataset!

```{r check-max}
met_avg <- met[, .(
  temp          = mean(temp,na.rm=T),
  rh            = mean(rh,na.rm=T),
  wind.sp       = mean(wind.sp,na.rm=T),
  vis.dist      = mean(vis.dist,na.rm=T),
  dew.point     = mean(dew.point,na.rm=T),
  lat           = mean(lat,na.rm=T),
  lon           = mean(lon,na.rm=T),
  elev          = mean(elev,na.rm=T)
), by = "USAFID"]
```

Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
```{r}
met_avg[, region:= fifelse(lon >= -98 & lat > 39.71, "NE",
               fifelse(lon < -98 & lat > 39.71, "NW",
               fifelse(lon < -98 & lat <= 39.71, 
"SW", "SE")))
      ]
table(met_avg$region)
```


Create a categorical variable for elevation as in the lecture slides

```{r}
met_avg[, elev_cat := fifelse(elev > 252, "high", "low")]
```


## Step 3. Make Violin plots of dew point temp by region

```{r violin-dewpoint}
met_avg[!is.na(region)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(x = 1, y = dew.point, color=region, fill = region)) + 
  facet_wrap(~ region, nrow = 1)
```

The highest dew point temperatures are reported in the southeast.


```{r violin-wind.sp}
met_avg[!is.na(region) & !is.na(wind.sp)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(x = 1, y = wind.sp, color=region, fill = region)) + 
  facet_wrap(~ region, nrow = 1)
```

The highest wind speeds are reported in the Southwest, but overall, there is not a huge difference between the regions.

## Step 4. Use geom_point with geom_smooth to examine the association between dew point temperature and wind speed by region

Colour points by region
Make sure to deal with NA category
Fit a linear regression line by region
Describe what you observe in the graph

```{r scatterplot-dewpoint-wind.sp}
met_avg[!is.na(region) & !is.na(wind.sp)] %>% 
  ggplot(mapping = aes(x = wind.sp , y = dew.point)) + 
  geom_point(mapping = aes(color = region)) +
  geom_smooth(method = lm, mapping = aes(linetype = region)) +
  facet_wrap(~ region, nrow = 1)
```

The southwest region shows a pattern of decrease while the other three regions are increasing.

## Step 5. Use geom_bar to create barplots of the weather stations by elevation category coloured by region

```{r}
met_avg %>%
filter(!(elev_cat %in% NA)) %>% 
  ggplot() + 
  geom_bar(mapping = aes(x = elev_cat, fill=region), position = "dodge") + scale_fill_brewer(palette = "Blues") +
  labs(title = "Elevation of Weather Stations") + 
  labs(x = expression("Elevation" * (degrees)), y = "Number of Weather Stations")
```


##Step 6. Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars

```{r}
met_avg[!is.na(dew.point)] %>%
  ggplot(mapping = aes(x=region, y=dew.point)) + 
    stat_summary(fun.data = mean_sdl, 
  geom = "errorbar")
```








