---
title: "lab06"
author: "Kaila An"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-libraries}
library(tidyverse)
library(tidytext)
library(dtplyr)
library(dplyr)
library(ggplot2)
library(forcats)
```


## Step 1. Read in the data
First download and then read in with read.csv()

```{r read-data, cache=TRUE}
if (!file.exists("mtsamples.csv"))
  download.file(
    url      = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv",
    destfile = "mtsamples.csv",
    method   = "libcurl",
    timeout  = 60
    )
mts <- read.csv("mtsamples.csv")
str(mts)
mts <- as_tibble(mts)
mts
```


## Question 1. What specialties do we have?
We can use count() from dplyr to figure out how many different catagories do we have? Are these categories related? Overlapping? Evenly distributed?

```{r medical-specialties}
specialties <-
    mts %>%
    count(medical_specialty)

specialties %>%
  arrange(desc(n)) %>%
knitr::kable()
```

There are `r nrow(specialties)` medical specialties.

```{r barplot-of-specialty-counts}
specialties %>%
  top_n(10) %>%
    ggplot(aes(x = n, y = fct_reorder(medical_specialty,n))) +
    geom_col()
```

The distribution is not all uniform.


## Question 2
Tokenize the the words in the transcription column
Count the number of times each token appears
Visualize the top 20 most frequent words
Explain what we see from this result. Does it makes sense? What insights (if any) do we get?

```{r token-transcription, cache=TRUE}
mts %>%
  unnest_tokens(word, transcription) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = c("word")) %>% 
  # use regular expression to filter out numbers
  filter( !grepl(pattern="^[0-9]+$", x=word)) %>%
  top_n(20,n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

There are a lot of stop words here, non-specific to medical text.
We do see "patient"!


## Question 3.
Redo visualization but remove stopwords before
Bonus points if you remove numbers as well
What do we see know that we have removed stop words? Does it give us a better idea of what the text is about?






