---
title: "lab06"
author: "Kaila An"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-libraries}
library(tidyverse)
library(tidytext)
library(dtplyr)
library(dplyr)
library(ggplot2)
library(forcats)
```


## Read in the data
First download and then read in with read.csv()

```{r read-data, cache=TRUE}
if (!file.exists("mtsamples.csv"))
  download.file(
    url      = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv",
    destfile = "mtsamples.csv",
    method   = "libcurl",
    timeout  = 60
    )
mts <- read.csv("mtsamples.csv")
str(mts)
mts <- as_tibble(mts)
mts
```


## Question 1. What specialties do we have?
We can use count() from dplyr to figure out how many different catagories do we have? Are these categories related? Overlapping? Evenly distributed?

```{r medical-specialties}
specialties <-
    mts %>%
    count(medical_specialty)

specialties %>%
  arrange(desc(n)) %>%
knitr::kable()
```

There are `r nrow(specialties)` medical specialties.

```{r barplot-of-specialty-counts}
specialties %>%
  top_n(10) %>%
    ggplot(aes(x = n, y = fct_reorder(medical_specialty,n))) +
    geom_col()
```

The distribution is not all uniform.


## Question 2
Tokenize the the words in the transcription column
Count the number of times each token appears
Visualize the top 20 most frequent words
Explain what we see from this result. Does it makes sense? What insights (if any) do we get?

```{r token-transcription, cache=TRUE}
mts %>%
  unnest_tokens(word, transcription) %>%
  #anti_join(stop_words, by = c("word")) %>%
  count(word, sort = TRUE) %>%
  top_n(20,n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

There are a lot of stop words here, non-specific to medical text.
We do see "patient"!


## Question 3.
Redo visualization but remove stop words before
Bonus points if you remove numbers as well

```{r, cache=TRUE}
mts %>%
  unnest_tokens(word, transcription) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = c("word")) %>% 
  # use regular expression to filter out numbers
  filter( !grepl(pattern="^[0-9]+$", x=word)) %>%
  top_n(20,n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

What do we see know that we have removed stop words? Does it give us a better idea of what the text is about?

  We see more medical terminology now that we have removed the stop words.  It gives us a better idea of what   the text is about.


## Question 4. Repeat question 2, but this time tokenize into bi-grams. How does the result change if you look at tri-grams?

```{r bigram-transcription, cache=TRUE}
mts %>%
  unnest_ngrams(bigram, transcription, n=2) %>%
  count(bigram, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(bigram, n))) +
  geom_col()
```

```{r trigrams-transcription, cache=TRUE}
mts %>%
  unnest_ngrams(trigram, transcription, n=3) %>%
  count(trigram, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(trigram, n))) +
  geom_col()
```

Top 20 trigrams were a bit more detailed than the bigrams and therefore, contained more medical terminology.


## Question 5. Using the results you got from questions 4. Pick a word and count the words that appears after and before it.

```{r bigrams-transcription, cache=TRUE}
ptbigram <-
mts %>%
  unnest_ngrams(bigram, transcription, n=2) %>%
  separate(bigram, into = c("word1", "word2"), sep = " ") %>%
  select(word1, word2) %>%
  filter(word1 == "patient" | word2 == "patient") 
```

```{r before-patient}
ptbigram %>%
  filter(word2=="patient") %>%
  count(word1, sort=TRUE) %>%
  anti_join(stop_words, by = c("word1" = "word")) %>%
  top_n(10) %>%
knitr::kable()
```

```{r after-patient}
ptbigram %>%
  filter(word1=="patient") %>%
  count(word2, sort=TRUE) %>%
  anti_join(stop_words, by = c("word2" = "word")) %>%
  top_n(10) %>%
knitr::kable()
```


## Question 6. Which words are most used in each of the specialties. You can use group_by() and top_n() from dplyr to have the calculations be done within each specialty. Remember to remove stopwords. How about the most 5 used words?

```{r top5-words-per-specialty, cache=TRUE}
mts %>%
  unnest_tokens(word, transcription) %>%
  group_by(medical_specialty) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = c("word")) %>%
  filter( !(word %in% stop_words$word) & !grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(5, n) %>%
  arrange(medical_specialty, desc(n)) %>%
knitr::kable()
```



